### 2022-03-18 paper/localpsf.tex
restructuring paper

%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
Our work builds upon a history of ``product-convolution`` approximations---weighted sums of convolution operators with spatially varying weights. A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators. Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}. Often product-convolution methods are known as ``point spread function'' (PSF) methods, particularly in applied fields such as optics and image processing. There, the ``operator'' $\Aop$ is a physical process of distortion and blurring that occurs as rays of light pass through an optical system like a telescope or microscope. However, ``PSF'' is also sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
For background on product-convolution and PSF methods, we recommend reading the following papers: \cite{PRODCONVGOOD}.

We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate approximation. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many operator actions of $\Aop$ are required.

In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$. This requires us to estimate the supports of the impulse responses, $\phi_x$, \emph{before} we compute them. We were inspired by resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$. If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions. Our method estimates the support of $\phi_x$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}. The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}. However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform. Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}. Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.

>>>>

%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
Product-convolution approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights. A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators. Our work, while not exactly a product-convolution approximation, builds upon a history of product-convolution approximations, which have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}. Often product-convolution methods are known as ``point spread function'' (PSF) methods, particularly in applied fields such as optics and image processing. There, the ``operator'' $\Aop$ is a physical process of distortion and blurring that occurs as rays of light pass through an optical system like a telescope or microscope. However, ``PSF'' is also sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
For background on product-convolution and PSF methods, we recommend reading the following papers: \cite{PRODCONVGOOD}.

We specifically build upon a class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate approximation. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many points $x_i$, and hence operator actions of $\Aop$, are required.

In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$. This requires us to estimate the supports of the impulse responses, $\phi_x$, \emph{before} we compute them. We were inspired to do this by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$. If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}). Our method estimates the support of $\phi_x$ more accurately and reliably than the techniques used in resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the techniques from resolution analysis do not have this requirement.
%(and moreover, the seismic Hessian does not have a non-negative integral kernel).

Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, allowing product-convolution approximations to be used as a fast surrogates for $\Aop$. But often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator. In this case, one can perform the necessary linear algebra operations using Krylov methods that only use matrix-vector products, replacing $\Aop$ with its product-convolution approximation \cite{PRODCONVKRYLOV}. The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$. However, there are several problems with this approach. First, the number of Krylov iterations can be large, particularly when $\Aop$ is high-rank. Second, complex geometries problems are rarely discretized using regular grids, but the FFT can only be used for functions that are sampled on a regular grid. Hence, using the FFT to perform convolutions requires transferring discretized functions from an irregular mesh to a regular grid and back. Third, the number of FFTs that must be performed is proportional to the number of convolution kernels. If a large number of convolution kernels are used, this approach can be slow.
%Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to product-convolution approximation are often sparse in the wavelet domain. Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.

Since our end goal is constructing a H-matrix, we are freed from strict adherence to the product-convolution format. This allows us to modify the product-convolution approach to address well-known deficiencies. In particular, product-convolution approximations are based on an assumption of local translation invariance, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}). Also, product-convolution approximations have a well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses. The product-convolution method requires one to extend the impulse responses outside of the domain where they are undefined. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}. In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined. In exchange for these benefits, our approximation can no longer be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT. But this is fine because our approximation is an H-matrix, which can be applied to vectors even faster.
